---
title: "synergy analysis"
output: html_notebook
---

```{r}
rm(list=ls())
experiment.title="H1s_raw"
```


## Load packages
```{r setup, include=FALSE}
###########
## Setup ##
###########

pacman::p_load(limma, edgeR, pheatmap, RColorBrewer, ggplot2, ggpubr, qvalue, plyr, wesanderson, GSEABase, grid, scales, WebGestaltR, stringr)
setwd("/Users/sadykom/Desktop/synergy/")
#wplot_save_this()
```

## custom functions
```{r}
mds <- function(normDGE, metacol, title){
  mcol <- as.factor(metacol)
  col <- rainbow(length(levels(mcol)), 1, 0.8, alpha = 0.5)[mcol]
  plotMDS(normDGE, col = col, pch = 16, cex = 2)
  legend("center", 
         fill = rainbow(length(levels(mcol)), 1, 0.8), 
         legend = levels(mcol), 
         horiz = F, 
         bty = "o", 
         box.col="grey", 
         xpd=TRUE)
  title(main=title)
}
```

```{r}
cameraplusplots <- function(contrast, genesetlist, vobject, design, catcolors, title){ 
  tmp.list <- list()
  cam <- data.frame(matrix(ncol = 5, nrow = 0))
  for (i in 1:length(genesetlist)){ 
    cam.s <- camera(vobject, genesetlist[[i]], design, contrast = contrast, inter.gene.cor = 0.01) 
    tmp.list[[i]] <- cam.s
    names(tmp.list)[i] <- names(genesetlist)[i] 
    tmp.list[[i]]$category <- names(tmp.list[i]) 
    colnames(cam) <- names(tmp.list[[1]]) 
    cam <- rbind.data.frame(cam, tmp.list[[i]]) 
    print(paste0("Gene set categories run: ", i))
  }
  cam$neglogFDR <- -log10(cam$FDR) 
  ## for plotting purposes only: 
  cam$dirNeglogFDR <- cam$neglogFDR
  cam[(cam$Direction == "Down"), "dirNeglogFDR"] <- -cam[(cam$Direction == "Down"), "neglogFDR"]
  grob <- grobTree(textGrob(c("UP","DOWN"), x = c(0.94, 0.89), y = c(0.95, 0.05), hjust = 0, gp = gpar(fontsize = 13))) 
  q <- ggplot(aes(x = cam$category, y = dirNeglogFDR, color = category), data = cam) +
    scale_color_manual(values = catcolors) +
    geom_jitter(aes(size = NGenes, alpha = neglogFDR), pch = 19, show.legend = F) +
    scale_size_continuous(range = c(4,16)) +
    scale_alpha_continuous(range = c(0.4, 1)) +
    geom_hline(yintercept = c(-1.3, 1.3), color = "red", alpha = 0.5) +
    geom_hline(yintercept = 0) +
    scale_y_continuous(limits = c(-10, 10), oob = squish, labels = abs) +
    labs(x = "Gene set categories", y = "-log10(FDR)", title = title) +
    theme_bw(14) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          axis.ticks.x = element_blank(),
          panel.grid.minor = element_blank(),
          panel.grid.major = element_blank()) +
    annotation_custom(grob) 
  print(q)
  cam$geneSet <- row.names(cam)
  cam10 <- as.data.frame(cam)
  cam10 <- cam10[order(cam10$FDR),]
  cam10 <- cam10[1:10,]
  grob <- grobTree(textGrob(c("DOWN","UP"), x = c(0.03, 0.9),  y=c(0.025), hjust = 0, gp = gpar(fontsize = 9, col = "grey60")))
  g <- ggplot(aes(x = geneSet, y = dirNeglogFDR, fill = category), data = cam10) +
    geom_col()+
    aes(reorder(stringr::str_wrap(geneSet, 60),-FDR), dirNeglogFDR) +
    xlab(NULL) +
    geom_hline(yintercept = c(-1.3, 1.3), color = "red", alpha = 0.3) +
    geom_hline(yintercept = 0) +
    scale_y_continuous(limits = c(-10, 10), oob = squish, labels = abs) +
    labs(y = "-log10(FDR)", title = title) +
    scale_fill_manual(values = catcolors) +
    coord_flip() +
    theme_bw() +
    theme(panel.grid.minor = element_blank(),
          panel.grid.major.y = element_blank()) +
    annotation_custom(grob) 
  print(g) 
  return(cam)
}
```

```{r}
power.compare.logFC <- function( sig1, sig2, N, N_other = c(2,4,6,8,10), alpha = 0.05, n_tests = 20000){
  d <- seq(0, 3, length.out=1000)
  alpha_multiple <- alpha / n_tests
  df <- lapply( N_other/N, function(n_scale){
    sigSq <- (sig1^2 + sig2^2) / n_scale
    cutoff <- qnorm( alpha_multiple/2, 0, sd = sqrt(sigSq), lower.tail = FALSE)
    p1 <- pnorm(-1*cutoff, d, sqrt(sigSq))
    p2 <- 1-pnorm(cutoff, d, sqrt(sigSq))
    data.frame(n_scale, d, power=p1+p2)
  })
  df <- do.call("rbind", df)
  ggplot(df, aes(d, power, color = as.factor(n_scale*N))) + 
    geom_line() + 
    theme_bw(14) + 
    theme(aspect.ratio = 1, plot.title = element_text(hjust = 0.5)) + 
    ylim(0, 1) + 
    scale_color_discrete("Samples") + 
    xlab(bquote(abs(logFC[observed] - logFC[expected]))) + 
    ggtitle("Power versus difference in logFC")
}
```

```{r}
categorize.synergy <- function(logFCmatrix, meanSE){
  m <- logFCmatrix
  m$magnitude.syn <- NA
  for (i in 1:length(m$Gene_name)){
    if (m$Synergistic.logFC[i] > meanSE){
      if (m$Additive.logFC[i] < -meanSE){
        if (m$Combinatorial.logFC[i] > meanSE){
          m$magnitude.syn[i] = "more.up"
        } else m$magnitude.syn[i] = "less.down"
      } else m$magnitude.syn[i] = "more.up"
    }
    else if (m$Synergistic.logFC[i] < -meanSE){
      if (m$Additive.logFC[i] > meanSE){
        if (m$Combinatorial.logFC[i] < -meanSE){
          m$magnitude.syn[i] = "more.down"
        } else m$magnitude.syn[i] = "less.up"
      } else m$magnitude.syn[i] = "more.down"
    } else m$magnitude.syn[i] = "same"
  }
  m$magnitude.syn <- as.factor(m$magnitude.syn)
  return(m)
}
```

```{r}
stratify.by.syn.cat <- function(log2FC.matrix.sub){
  synergy.cat.list <- list("less.down" = as.character(log2FC.matrix.sub[
    log2FC.matrix.sub$magnitude.syn == "less.down", "Gene_name"]),
    "less.up" = as.character(log2FC.matrix.sub[
      log2FC.matrix.sub$magnitude.syn == "less.up", "Gene_name"]),
    "more.down" = as.character(log2FC.matrix.sub[
      log2FC.matrix.sub$magnitude.syn == "more.down", "Gene_name"]),
    "more.up" = as.character(log2FC.matrix.sub[
      log2FC.matrix.sub$magnitude.syn == "more.up", "Gene_name"]),
    "same" = as.character(log2FC.matrix.sub[
      log2FC.matrix.sub$magnitude.syn == "same", "Gene_name"]))
  return(synergy.cat.list)
}
```

## Load data
```{r}
meta <- read.csv("meta_all_noOutl.csv", row.names = 1)
counts <- read.csv("batch_corrected_SKO_DKO_noOutl_minCount1.csv", row.names=1)
meta <- meta[match(colnames(counts), row.names(meta)),]
```

## Filter low expr genes

```{r}
#pdf(paste0("results/", experiment.title, "-1_cpm-counts.pdf"))
plot(cpm(counts)[, 1], counts[, 1], ylim = c(0, 50), xlim = c(0, 3))
abline(h = 10, col = "red")
abline(v = 0.17, col = "red")
```


```{r}
#dev.off()
keep <- rowSums(cpm(counts[]) > 0.17) >= 100
gExpr <- counts[keep, ]
dim(gExpr)
```
## DGE list
```{r}
y <- DGEList(gExpr)
y <- calcNormFactors(y)
anno <- read.csv("anno.csv")
row.names(anno) <- anno$ensembl
anno <- anno[match(rownames(y), rownames(anno)), ]
y$genes <- anno
```

## Diagnositic Plots
```{r}
pdf(paste0("results/", experiment.title, "-2_mds_bc.pdf"))
for (i in 1:length(colnames(meta))){
  mds(y, meta[ ,i], colnames(meta)[i])
}
```

```{r}
plotMDS(y)
```
## FITTING LINEAR MODEL 
```{r}
design <- model.matrix(~ 0 + Tx, meta)
colnames(design) <- gsub("Tx", "", colnames(design))
#### Optional: Subsequently remove the column name from the created design matrix to ease its use.
colnames(design) <- gsub("BatchBatch", "Batch", colnames(design))
colnames(design)
```

```{r}
v <- voom(y, design, plot = TRUE, save.plot = TRUE)
## Save the plot (optional)
pdf(paste0("results/", experiment.title, "-3_voom.pdf"))
plot(v$voom.xy, type = "p", pch=20, cex=0.16, 
     main = "voom: Mean-variance trend", 
     xlab = "log2( count size + 0.5 )", 
     ylab = "Sqrt( standard deviation )")
lines(v$voom.line, col="red")
dev.off()
```
```{r}
fit <- lmFit(v, design)
```
## Make contrasts
```{r}
cont.matrix <- makeContrasts(
  dk0_1vsWT = DKO0_1 - WT,
  dk0_2vsWT = DKO0_2 - WT,
  dk0_3vsWT = DKO0_3 - WT,
  dk0_4vsWT = DKO0_4 - WT,
  dk0_5vsWT = DKO0_5 - WT,

  dk1_2vsWT = DKO1_2 - WT,
  dk1_4vsWT = DKO1_4 - WT,
  dk1_5vsWT = DKO1_5 - WT,

  dk2_3vsWT = DKO2_3 - WT,
  dk2_4vsWT = DKO2_4 - WT,
  dk2_5vsWT = DKO2_5 - WT,

  dk3_1vsWT = DKO3_1 - WT,
  dk3_4vsWT = DKO3_4 - WT,
  dk3_5vsWT = DKO3_5 - WT,

  dk4_5vsWT = DKO4_5 - WT,

  dkX_0vsWT = DKOX_0 - WT,
  dkX_1vsWT = DKOX_1 - WT,
  dkX_2vsWT = DKOX_2 - WT,
  dkX_3vsWT = DKOX_3 - WT,
  dkX_4vsWT = DKOX_4 - WT,
  dkX_5vsWT = DKOX_5 - WT,
  
  H1.0vWT = SKO_0 - WT,
  H1.1vWT = SKO_1 - WT,
  H1.2vWT = SKO_2 - WT,
  H1.3vWT = SKO_3 - WT,
  H1.4vWT = SKO_4 - WT,
  H1.5vWT = SKO_5 - WT,
  H1.XvWT = SKO_X - WT,
  
  AddSKO_0plus1 = SKO_0 + SKO_1 - 2*WT, # expectation of depletion
  SynSKO_0plus1 = (DKO0_1 - WT) - (SKO_0 - WT) - (SKO_1 - WT),
  
  AddSKO_0plus2 = SKO_0 + SKO_2 - 2*WT, # expectation of depletion
  SynSKO_0plus2 = (DKO0_2 - WT) - (SKO_0 - WT) - (SKO_2 - WT),
  
  AddSKO_0plus3 = SKO_0 + SKO_3 - 2*WT, # expectation of depletion
  SynSKO_0plus3 = (DKO0_3 - WT) - (SKO_0 - WT) - (SKO_3 - WT),
  
  AddSKO_0plus4 = SKO_0 + SKO_4 - 2*WT, # expectation of depletion
  SynSKO_0plus4 = (DKO0_4 - WT) - (SKO_0 - WT) - (SKO_4 - WT),
  
  AddSKO_0plus5 = SKO_0 + SKO_5 - 2*WT, # expectation of depletion
  SynSKO_0plus5 = (DKO0_5 - WT) - (SKO_0 - WT) - (SKO_5 - WT),
  
  AddSKO_1plus2 = SKO_1 + SKO_2 - 2*WT, # expectation of depletion
  SynSKO_1plus2 = (DKO1_2 - WT) - (SKO_1 - WT) - (SKO_2 - WT),
  
  AddSKO_1plus4 = SKO_1 + SKO_4 - 2*WT, # expectation of depletion
  SynSKO_1plus4 = (DKO1_4 - WT) - (SKO_1 - WT) - (SKO_4 - WT),
  
  AddSKO_1plus5 = SKO_1 + SKO_5 - 2*WT, # expectation of depletion
  SynSKO_1plus5 = (DKO1_5 - WT) - (SKO_1 - WT) - (SKO_5 - WT),
  
  AddSKO_2plus3 = SKO_2 + SKO_3 - 2*WT, # expectation of depletion
  SynSKO_2plus3 = (DKO2_3 - WT) - (SKO_2 - WT) - (SKO_3 - WT),
  
  AddSKO_2plus4 = SKO_2 + SKO_4 - 2*WT, # expectation of depletion
  SynSKO_2plus4 = (DKO2_4 - WT) - (SKO_2 - WT) - (SKO_4 - WT),
  
  AddSKO_2plus5 = SKO_2 + SKO_5 - 2*WT, # expectation of depletion
  SynSKO_2plus5 = (DKO2_5 - WT) - (SKO_2 - WT) - (SKO_5 - WT),
  
  AddSKO_3plus1 = SKO_1 + SKO_3 - 2*WT, # expectation of depletion
  SynSKO_3plus1 = (DKO3_1 - WT) - (SKO_3 - WT) - (SKO_1 - WT),
  
  AddSKO_3plus4 = SKO_4 + SKO_3 - 2*WT, # expectation of depletion
  SynSKO_3plus4 = (DKO3_4 - WT) - (SKO_3 - WT) - (SKO_4 - WT),
  
  AddSKO_3plus5 = SKO_5 + SKO_3 - 2*WT, # expectation of depletion
  SynSKO_3plus5 = (DKO3_5 - WT) - (SKO_3 - WT) - (SKO_5 - WT),
  
  AddSKO_4plus5 = SKO_5 + SKO_4 - 2*WT, # expectation of depletion
  SynSKO_4plus5 = (DKO4_5 - WT) - (SKO_4 - WT) - (SKO_5 - WT),
  
  AddSKO_Xplus0 = SKO_X + SKO_0 - 2*WT, # expectation of depletion
  SynSKO_Xplus0 = (DKOX_0 - WT) - (SKO_X - WT) - (SKO_0 - WT),
  
  AddSKO_Xplus1 = SKO_X + SKO_1 - 2*WT, # expectation of depletion
  SynSKO_Xplus1 = (DKOX_1 - WT) - (SKO_X - WT) - (SKO_1 - WT),
  
  AddSKO_Xplus2 = SKO_X + SKO_2 - 2*WT, # expectation of depletion
  SynSKO_Xplus2 = (DKOX_2 - WT) - (SKO_X - WT) - (SKO_2 - WT),
  
  AddSKO_Xplus3 = SKO_X + SKO_3 - 2*WT, # expectation of depletion
  SynSKO_Xplus3 = (DKOX_3 - WT) - (SKO_X - WT) - (SKO_3 - WT),
  
  AddSKO_Xplus4 = SKO_X + SKO_4 - 2*WT, # expectation of depletion
  SynSKO_Xplus4 = (DKOX_4 - WT) - (SKO_X - WT) - (SKO_4 - WT),
  
  AddSKO_Xplus5 = SKO_X + SKO_5 - 2*WT, # expectation of depletion
  SynSKO_Xplus5 = (DKOX_5 - WT) - (SKO_X - WT) - (SKO_5 - WT),
  
  levels = design)
```
## Visualize the contrasts in a heatmap
```{r}
cont.p <- t(cont.matrix)
h <- pheatmap(cont.p,
              display_numbers = T, number_format = "%.0f",
              breaks = seq(-3, 1, by = 0.5),
              color = colorRampPalette(rev(brewer.pal(n = 10, name = "RdYlBu")))(12), 
              cluster_cols = F, cluster_rows = F)
print(h)
```

## Assessing differential expression
```{r}
fit.cont <- contrasts.fit(fit, cont.matrix)
fit.cont <- eBayes(fit.cont)
plotSA(fit.cont, main = "Final model: Mean-variance trend", ylab = "Sqrt( standard deviation )")

```

```{r}
summa.fit = decideTests(fit.cont, adjust.method = "fdr")
```
##  Save DEG result tables
```{r}
res.list <- list()
for (i in 1:length(colnames(fit.cont$contrasts))){
  x <- topTable(fit.cont, coef = i, sort.by = "p", n = Inf, confint = T)
  res.list[[i]] <- x
  names(res.list)[i] <- colnames(fit.cont$contrasts)[i]
  write.csv(x, paste0("results/", experiment.title, "_DEGs_", 
                      colnames(fit.cont$contrasts)[i], ".csv"))
}
#### Optional: Create separate variable for each list entry.
list2env(res.list, globalenv())
```

## Save DEG result plots
```{r}
pdf(paste0("results/", experiment.title, "-4_volcano-md-plots.pdf"))
par(mfrow = c(1, 2))
for (i in 1:length(colnames(fit.cont$contrasts))){
  plotMD(fit.cont, coef = i, status = summa.fit[, i], values = c(-1, 1))
  volcanoplot(fit.cont, coef = i, highlight = 10, 
              names = fit.cont$genes$Gene_name)
}
```
## Top 3 DEG optional
```{r}
pdf(paste0("results/", experiment.title, "-5_top3-expression-plots.pdf"))
par(mfrow = c(1, 1))
for (i in 1:length(colnames(fit.cont$contrasts))){
  x <- topTable(fit.cont, coef = i, sort.by = "p", n = Inf)
  cat("  \n\n### Plotting",  colnames(fit.cont$contrasts)[i], "  \n\n")
  for (j in 1:3){
    deg <- as.character(x[j,"ensembl"])
    p <- qplot(meta$Tx, v$E[deg, ], 
               geom = "boxplot", fill = meta$Tx, 
               ylab = "Normalized expression", xlab = "group", 
               main = paste0(j, ". DEG: ", as.character(x[j, "Gene_name"]))) +
      geom_jitter() +
      #scale_x_discrete(limits = c("ctrl", "sanp91", "tsnare1", "clcn3", "furin", "all.ctrl", "all")) +
      #scale_fill_manual(values = (c("orchid4", "grey", "steelblue", "grey", "firebrick", "blue", "darkblue"))) +
      rotate_x_text(angle = 45) +
      theme_bw(14)+
      theme(legend.position = "none", 
            axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
            print(p)
  }
}
dev.off()
```

## Determining power to detect synergistic effects

```{r}
SE <- sqrt(fit.cont$s2.post) * fit.cont$stdev.unscaled
#### Calculate power. Choose the SE matrix column names that represent the additive and the combinatorial perturbation to calculate the median standard error. Then use them to run the power.compare.logFC() function, which creates a power plot (Fig. 6A).
colnames(SE)
```

```{r}
sig1 <- median(SE[,"AddSKO_0plus1"])
sig2 <- median(SE[,"dk0_1vsWT"])
g <- power.compare.logFC(sig1, sig2, N = 4, N_other = c(4, 6, 8, 10, 14), 
                         alpha = 0.05, n_tests = 20000)
pdf(paste0("results/", experiment.title, "-6_synergy-power_0_1.pdf"))
print(g)
dev.off()
```
## Determining the extent of synergy
```{r}
syn_list = c("SynSKO_0plus1", "SynSKO_0plus2", "SynSKO_0plus3", "SynSKO_0plus4", "SynSKO_0plus5", "SynSKO_1plus2", "SynSKO_1plus4", "SynSKO_1plus5", "SynSKO_2plus3", "SynSKO_2plus4", "SynSKO_2plus5", "SynSKO_3plus1", "SynSKO_3plus4", "SynSKO_3plus5", "SynSKO_4plus5", "SynSKO_Xplus0", "SynSKO_Xplus1", "SynSKO_Xplus2", "SynSKO_Xplus3", "SynSKO_Xplus4", "SynSKO_Xplus5")
add_list = c("AddSKO_0plus1", "AddSKO_0plus2", "AddSKO_0plus3", "AddSKO_0plus4", "AddSKO_0plus5", "AddSKO_1plus2", "AddSKO_1plus4", "AddSKO_1plus5", "AddSKO_2plus3", "AddSKO_2plus4", "AddSKO_2plus5", "AddSKO_3plus1", "AddSKO_3plus4", "AddSKO_3plus5", "AddSKO_4plus5", "AddSKO_Xplus0", "AddSKO_Xplus1", "AddSKO_Xplus2", "AddSKO_Xplus3", "AddSKO_Xplus4", "AddSKO_Xplus5")
com_list = c("dk0_1vsWT", "dk0_2vsWT", "dk0_3vsWT", "dk0_4vsWT", "dk0_5vsWT", "dk1_2vsWT", "dk1_4vsWT", "dk1_5vsWT", "dk2_3vsWT", "dk2_4vsWT", "dk2_5vsWT", "dk3_1vsWT", "dk3_4vsWT", "dk3_5vsWT", "dk4_5vsWT", "dkX_0vsWT", "dkX_1vsWT", "dkX_2vsWT", "dkX_3vsWT", "dkX_4vsWT", "dkX_5vsWT")
experiment.titles = c("SynSKO_0plus1", "SynSKO_0plus2", "SynSKO_0plus3", "SynSKO_0plus4", "SynSKO_0plus5", "SynSKO_1plus2", "SynSKO_1plus4", "SynSKO_1plus5", "SynSKO_2plus3", "SynSKO_2plus4", "SynSKO_2plus5", "SynSKO_3plus1", "SynSKO_3plus4", "SynSKO_3plus5", "SynSKO_4plus5", "SynSKO_Xplus0", "SynSKO_Xplus1", "SynSKO_Xplus2", "SynSKO_Xplus3", "SynSKO_Xplus4", "SynSKO_Xplus5")

# for (i in 1:length(my_list)) {
#   synergy.pvalues <- res.list[syn_list[i]][[1]][,"P.Value"]
#   print(synergy.pvalues[1])
#   add.pvalues <- res.list[add_list[i]][[1]][,"P.Value"]
#   print(add.pvalues[1])
#   print(experiment.titles[i])
# }


for (i in 1:length(syn_list)) {
#synergy.pvalues <- res.list$SynSKO_0plus1$P.Value
synergy.pvalues <- res.list[syn_list[i]][[1]][,"P.Value"]


pi1 <- 1 - qvalue(synergy.pvalues)$pi0
print(pi1) # 0.3624806 vs bn 0.7002615 (0_1) VS 0.8153062(3_1) or 0.9659436

pdf(paste0("results/", experiment.title, "-7_synergy-coefficient_", experiment.titles[i], ".pdf"))
plot.new()
text(0.4,0.75,labels=paste0("\n", experiment.titles[i], "\n", round(pi1 * 100, 2), 
                            "% non-null \np-values and \n", 
                            round(sum(res.list[syn_list[i]][[1]][,"adj.P.Val"] < 0.1) *
                                    100/length(res.list[syn_list[i]][[1]][,"ensembl"]), 2), 
                            " % of genes with \nsynergy FDR < 0.1"))
dev.off()
meanSE = mean(SE[,c(1,2,3,4,6)]) # 0.34 vs 0.1289881 (0_1) vs 0.1289881 (3_1)
#colnames(res.list$AddSKO_0plus1)
log2FC.matrix <- Reduce(function(x,y) merge(x,y, by=c("ensembl", "Gene_name"), all = TRUE),
                        list(res.list[add_list[i]][[1]][,c(1,2,4,9)],
                             res.list[com_list[i]][[1]][,c(1,2,4,9)],
                             res.list[syn_list[i]][[1]][,c(1,2,4,9)]))
colnames(log2FC.matrix)
colnames(log2FC.matrix) <- c("Ensembl", "Gene_name", 
                             "Additive.logFC", "Additive.FDR", 
                             "Combinatorial.logFC", "Combinatorial.FDR", 
                             "Synergistic.logFC", "Synergistic.FDR")
rownames(log2FC.matrix) <- log2FC.matrix$Ensembl
log2FC.matrix <- categorize.synergy(log2FC.matrix, meanSE)
write.csv(log2FC.matrix, paste0("results/", experiment.title, "_LogFC-FDR-synergy_matrix_", experiment.titles[i], ".csv"))
genes.per.category <- count(log2FC.matrix, "magnitude.syn")
print(genes.per.category)
genes.per.category$category <- factor(genes.per.category$magnitude.syn,
                                      levels=c("same", "less.up", "less.down", "more.up", "more.down"))
genes.per.category$percent <- paste0(round(genes.per.category$freq*100/sum(genes.per.category$freq), 0), " %")
write.csv(genes.per.category, paste0("results/", experiment.title, "_gene-count_synergy-categories_", experiment.titles[i], ".csv"))
zissou <- wes_palette("Zissou1", 6, type = "continuous")
q <- ggplot(genes.per.category, aes(x = "", y = freq, fill = category)) +
  geom_col() +
  coord_polar("y", start=0) + 
  scale_fill_manual(values=zissou) +
  theme_void()
pdf(paste0("results/", experiment.title, "-8_synergy-categories_pie-chart_", experiment.titles[i], ".pdf"))
print(q)
dev.off()

pdf(paste0("results/", experiment.title, "-9_heatmaps_log2FC-Add-vs-Combi_", experiment.titles[i], ".pdf"))
for (j in 1:length(levels(log2FC.matrix$magnitude.syn))){
  breaks <- c(seq(-6, -0.3,by=0.1),seq(0.3, 6,by=0.1))
  breaks <- append(breaks, -9,0)
  breaks <- append(breaks, 9)
  tmp <- log2FC.matrix[log2FC.matrix$magnitude.syn == 
                         levels(log2FC.matrix$magnitude.syn)[j],
                       c("Additive.logFC","Combinatorial.logFC")]
  h <- pheatmap(tmp,
                kmeans_k = 30, 
                cellwidth = 70, cellheight = 5,
                border_color = NA, 
                breaks=breaks,
                cluster_cols = F,
                show_rownames = F,
                color = colorRampPalette(rev(brewer.pal(n=9, name="RdBu")))(117),
                main = paste0("logFC expected vs. measured:\n",
                              levels(log2FC.matrix$magnitude.syn)[j]))
}
dev.off()
}
```
## Determining the extent of synergy
```{r}
pdf(paste0("results/", experiment.title, "-7_synergy-coefficient_3_1.pdf"))
plot.new()
text(0.4,0.75,labels=paste0("\n",round(pi1 * 100, 2), 
                            "% non-null \np-values and \n", 
                            round(sum(res.list$SynSKO_0plus1$adj.P.Val < 0.1) *
                                    100/length(res.list$SynSKO_0plus1$ensembl), 2), 
                            " % of genes with \nsynergy FDR < 0.1"))
```

```{r}
hist(synergy.pvalues)
```

```{r}
meanSE = mean(SE[,c(1,2,3,4,6)]) # 0.34 vs 0.1289881 (0_1) vs 0.1289881 (3_1)
colnames(res.list$AddSKO_0plus1)

```

```{r}
log2FC.matrix <- Reduce(function(x,y) merge(x,y, by=c("ensembl", "Gene_name"), all = TRUE),
                        list(res.list$AddSKO_0plus1[,c(1,2,4,9)],
                             res.list$dk0_1vsWT[,c(1,2,4,9)],
                             res.list$SynSKO_0plus1[,c(1,2,4,9)]))
colnames(log2FC.matrix)
```

```{r}
nrow(unique(res.list$dk0_1vsWT[,c("ensembl", "Gene_name")])) # 23275
nrow(res.list$dk0_1vsWT[,c("ensembl", "Gene_name")]) # 23275
```


```{r}
colnames(log2FC.matrix) <- c("Ensembl", "Gene_name", 
                             "Additive.logFC", "Additive.FDR", 
                             "Combinatorial.logFC", "Combinatorial.FDR", 
                             "Synergistic.logFC", "Synergistic.FDR")
rownames(log2FC.matrix) <- log2FC.matrix$Ensembl
```

```{r}
log2FC.matrix <- categorize.synergy(log2FC.matrix, meanSE)
write.csv(log2FC.matrix, paste0("results/", experiment.title, "_LogFC-FDR-synergy_matrix_3_1.csv"))
genes.per.category <- count(log2FC.matrix, "magnitude.syn")
print(genes.per.category)
```

```{r}
genes.per.category$category <- factor(genes.per.category$magnitude.syn,
                                      levels=c("same", "less.up", "less.down", "more.up", "more.down"))
genes.per.category$percent <- paste0(round(genes.per.category$freq*100/sum(genes.per.category$freq), 0), " %")
write.csv(genes.per.category, paste0("results/", experiment.title, "_gene-count_synergy-categories_3_1.csv"))
```

## Visualize categories
```{r}
zissou <- wes_palette("Zissou1", 6, type = "continuous")
q <- ggplot(genes.per.category, aes(x = "", y = freq, fill = category)) +
  geom_col() +
  coord_polar("y", start=0) + 
  scale_fill_manual(values=zissou) +
  theme_void()
pdf(paste0("results/", experiment.title, "-8_synergy-categories_pie-chart_0_1c.pdf"))
print(q)
```
## Visualize the categories in heatmaps
```{r}
pdf(paste0("results/", experiment.title, "-9_heatmaps_log2FC-Add-vs-Combi_3_1d.pdf"))
for (i in 1:length(levels(log2FC.matrix$magnitude.syn))){
  breaks <- c(seq(-6, -0.3,by=0.1),seq(0.3, 6,by=0.1))
  breaks <- append(breaks, -9,0)
  breaks <- append(breaks, 9)
  tmp <- log2FC.matrix[log2FC.matrix$magnitude.syn == 
                         levels(log2FC.matrix$magnitude.syn)[i],
                       c("Additive.logFC","Combinatorial.logFC")]
  h <- pheatmap(tmp,
                kmeans_k = 30, 
                cellwidth = 70, cellheight = 5,
                border_color = NA, 
                breaks=breaks,
                cluster_cols = F,
                show_rownames = F,
                color = colorRampPalette(rev(brewer.pal(n=9, name="RdBu")))(117),
                main = paste0("logFC expected vs. measured:\n",
                              levels(log2FC.matrix$magnitude.syn)[i]))
}
```















